---
title: "p8105_hw1_yp2743"
author: "Yueran Pan"
date: "2026-09-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

Description of penguins dataset:

Dataset 'penguins' contain `r nrow(penguins)` observations with `r ncol(penguins)` variables, and each row represents one penguin. Variables include species (`species`: Adelie, Chinstrap, and Gentoo), island (`island`:Biscoe, Dream, and Torgersen), bill depth (`bill_depth_mm`) and bill length (`bill_length_mm`) in millimeters, sex (`sex`), body mass (`body_mass_g`) in grams, and year (`year`).

The average flipper length among the studied penguin is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`mm.

## Scatterplot

Below is a scatterplot using bill length (x-axis) against flipper length (y-axis), with points colored by species. Two observations with missing measurements are omitted from the plot.

```{r prob1_ggplot}
ggplot(penguins, aes(x = bill_length_mm,  y = flipper_length_mm, color = species)) + 
    geom_point()
#Export using ggsave
ggsave("hw1_penguins_scatterplot.pdf", height = 4, width = 6)
```




# Problem 2

## Create a data frame

- a random sample of size 10 from a standard Normal distribution

- a logical vector indicating whether elements of the sample are greater than 0

- a character vector of length 10

- a factor vector of length 10, with 3 different factor “levels”

```{r prob2_dataframe}
hw1_df = tibble(
  sample = rnorm(10),
  logical_vec = sample > 0,
  char_vec = c("y","p", "2","7", "4", "3", "H", "W", "No.", "1"),
  factor_vec = factor(c("medium","grande","venti","medium","grande","venti","medium","grande","venti","medium"))
)

```

## Take the mean of a variable in a dataframe,


```{r prob2_mean, collapse = TRUE}
mean(pull(hw1_df, sample))
mean(pull(hw1_df, logical_vec))
mean(pull(hw1_df, char_vec))
mean(pull(hw1_df, factor_vec))
```
From `sample`, which is a numeric variable, I got a valid mean. 

From `logical_vec`, which is a logical vector generates TRUE or FALSE, I got mean of 0.4.

From `char_vec`, which is a character vector, I got NA.

From `factor_vec`, which is a character vector, I got NA.

## Code chunk convertlogical, character, and factor variables to numeric.
```{r prob2_as_numer, results = "hide", WARNING = FALSE}
as.numeric(pull(hw1_df, logical_vec))
as.numeric(pull(hw1_df, char_vec))
as.numeric(pull(hw1_df, factor_vec))
```


## What happens, and why? 
For logical vector, numeric conversion leads to 0 and 1 result, which is true or false. I also receive a warning from R that NAs introduced by coercion.

For character vector, numeric conversion leads to NA and number result, this makes sense because I treat my uni as characters, but four digits are actually number.

For level vector, numeric conversion leads to 1, 2, and 3. `medium` is level 1, `grande` is level 2, `venti` is level 3. This makes sense, R treats it into three level that has an hierarchy.


## Does this help explain what happens when you try to take the mean?
Yes, ~~~??