---
title: "p8105_hw1_yp2743"

author: "Yueran Pan"

date: "2026-09-26"

output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

Description of penguins dataset:

Dataset 'penguins' contain `r nrow(penguins)` observations with `r ncol(penguins)` variables, and each row represents one penguin. Variables include species (`species`: Adelie, Chinstrap, and Gentoo), island (`island`:Biscoe, Dream, and Torgersen), bill depth (`bill_depth_mm`) and bill length (`bill_length_mm`) in millimeters, sex (`sex`), body mass (`body_mass_g`) in grams, and year (`year`).

The average flipper length among the studied penguin is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`mm.

### Scatterplot

Below is a scatterplot using bill length (x-axis) against flipper length (y-axis), with points colored by species. Two observations with missing measurements are omitted from the plot.

```{r prob1_ggplot}
ggplot(penguins, aes(x = bill_length_mm,  y = flipper_length_mm, color = species)) + 
    geom_point()
#Export using ggsave
ggsave("hw1_penguins_scatterplot.pdf", height = 4, width = 6)
```




## Problem 2

### Create a data frame

- a random sample of size 10 from a standard Normal distribution

- a logical vector indicating whether elements of the sample are greater than 0

- a character vector of length 10

- a factor vector of length 10, with 3 different factor “levels”

```{r prob2_dataframe}
hw1_df = tibble(
  sample = rnorm(10),
  logical_vec = sample > 0,
  char_vec = c("y","p", "2","7", "4", "3", "H", "W", "No.", "1"),
  factor_vec = factor(c("medium","grande","venti","medium","grande","venti","medium","grande","venti","medium"))
)

```

### Take the mean of a variable in a dataframe


```{r prob2_mean, collapse = TRUE}
mean(pull(hw1_df, sample))
mean(pull(hw1_df, logical_vec))
mean(pull(hw1_df, char_vec))
mean(pull(hw1_df, factor_vec))
```


From `sample`, which is a numeric variable, I got a valid mean. 

From `logical_vec`, which is a logical vector generates TRUE or FALSE, I got valid mean.

From `char_vec`, which is a character vector, I got NA mean, with a warning.

From `factor_vec`, which is a factor vector, I got NA, with a warning.


### Code chunk convertlogical, character, and factor variables to numeric
```{r prob2_as_numer, results = "hide", WARNING = FALSE}
as.numeric(pull(hw1_df, logical_vec))
as.numeric(pull(hw1_df, char_vec))
as.numeric(pull(hw1_df, factor_vec))
```


### What happens, and why? 

For `logical_vec`, numeric conversion leads to 0 and 1 result, each recorded obs will be either 1 (originally, true) or 0 (originally, false). 

I receive a warning from R that `char_vec`(character vector)'s numeric conversion NAs introduced by coercion.


For char_vec`, numeric conversion leads to NA and number result. If I have put ten length all by letters, I would have gotten all 'NA'. But because I add strings "2", "7", "4", "3", and "1", these can be interpreted as numbers and thus became numbers. The remaining strings cannot be intrepreted as numbers became 'NA', which produces the warning "NAs introduced by coercion".


For `factor_vec`, numeric conversion leads to 1, 2, and 3. `medium` = 2, `grande` = 1, `venti` = 3. Instead of following first appearring became the first, my variables follow the alphabetical order of levels. Despite I wrote cup size level, but the factor vector I created is not an ordered factor.


### Does this help explain what happens when I try to take the mean?


These conversions help explain why same `mean()` function yield to different results, because I create four different types of vector. 

Numeric values can be averaged directly, and logical values can be also averaged, as `TRUE` or `FALSE` representing 0/1.

As `mean()` does not automatically convert characters to numbers, averaging character vector `char_vec` or factor vector `faactor_vec` would only yield into N/A result.



(p.s: If I treat numbers as characters, for example, zip code or my id, conversion function can convert them back to number that R can recognize (as numeric). However if strings were letter, each string would become an NA.
R can convert factor vector to numeric value, but the converted numbers lack meanings or loses the semantic meaning.) 