---
title: "p8105_hw1_yp2743"
author: "Yueran Pan"
date: "2026-09-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")

```

Description of penguins dataset:

Dataset 'penguins' contain `r nrow(penguins)` observations with `r ncol(penguins)` variables, and each row represents one penguin. Variables include species (`species`: Adelie, Chinstrap, and Gentoo), island (`island`:Biscoe, Dream, and Torgersen), bill depth (`bill_depth_mm`) and bill length (`bill_length_mm`) in millimeters, sex (`sex`), body mass (`body_mass_g`) in grams, and year (`year`).

The average flipper length among the studied penguin is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`mm.

# Scatterplot

Below is a scatterplot using bill length (x-axis) against flipper length (y-axis), with points colored by species. Two observations with missing measurements are omitted from the plot.

```{r}
ggplot(penguins, aes(x = bill_length_mm,  y = flipper_length_mm, color = species)) + 
    geom_point()
#Export using ggsave
ggsave("hw1_penguins_scatterplot.pdf", height = 4, width = 6)
```




## Problem 2